---
title: "Week 2 Report"
author: "Diana Hilleshein"
date: "1/24/2022"
output: pdf_document

---
## General information on the data  
Data was downloaded from: <https://www.europeansocialsurvey.org/download.html?file=ESS9e03_1&y=2018>.    
Codebook fro the data can be found by the following link: <chrome-extension://efaidnbmnnnibpcajpcglclefindmkaj/viewer.html?pdfurl=https%3A%2F%2Fwww.europeansocialsurvey.org%2Fdocs%2Fround9%2Fsurvey%2FESS9_appendix_a7_e03_1.pdf&clen=2111156>.  

The European Social Survey (ESS) is an academically-driven multi-country survey, which has been administered in over 35 countries to date. Its three aims are, firstly – to monitor and interpret changing public attitudes and values within Europe and to investigate how they interact with Europe's changing institutions, secondly – to advance and consolidate improved methods of cross-national survey measurement in Europe and beyond, and thirdly – to develop a series of European social indicators, including attitudinal indicators. For the course we will use round 9 ESS.

Variables of the datases are separated for the following groups:   
 Country  
 Weights  
 Media and social trust  
 Politics  
 Subjective well-being, social exclusion, religion, national and ethnic identity  
 Timing of life  
 Gender, Year of birth and Household grid  
 Socio-demographics  
 Justice and Fairness  
 Human values  
 Administrative variables  
 Sample Design variables  
 User defined variables  
 
```{r include=FALSE}
setwd("C:/R scripts/DWA2022")

# load/attach package 
library(haven)

# import data
data <- read_sav("C:/R scripts/DWA2022/ESS9e03_1.sav")


View(data)

save(data,file="data.Rda")
```

## Part 1. Explore the European Social Survey (ESS) data.  
I am interested in the following self-made groups of variables:  
*Involovemt in politics*  
1) *polintr* -How interested in politics (scale from 1 to 4, 1= Very interested, 4 = Not at all interested)  
2) *actrolga* -Able to take active role in political group (scale from 1 to 5, 1 = Not at all able, 5 = Completely able)  
*Attitude to immigrants*  
3) *imwbcnt* -Immigrants make country worse or better place to live (scale from 0 to 10, 0 = Worse place to live, 10 = Better place to live)  
4) *imbgeco* -Immigration bad or good for country's economy  (scake from 0 to 10, 0 = Bad for the economy, 10 = Good for the economy)
5) *imueclt* -Country's cultural life undermined or enriched by immigrants (scale from 0 to 10, 0 = Cultural life undermined, 10 = Cultural life enriched)  
6) *imsmetn* -Allow many/few immigrants of same race/ethnic group as majority (scale from 1 to 4, 1 = Allow many to come and live here, 4 = Allow none)  
7) *imdfetn* -Allow many/few immigrants of different race/ethnic group from majority (scale from 1 to 4, 1 = Allow many to come and live here, 4 = Allow none)  
8) *impcntr* -Allow many/few immigrants from poorer countries outside Europe (scale from 1 to 4, 1 = Allow many to come and live here, 4 = Allow none)  
*General information on a participants*  
9) *stflife* -How satisfied with life as a whole (from 0 to 10, 0 = extremely dissatisfied, 10 = extremely satisfied)  
10) *gndr*-Gender (1 = Male, 2 = Female)  
11) *agea* -Age of respondent, calculated (integer)  

Comments on the variables: 
- variables polintr and actrolga need to be recoded since they have opposite meanings for the lowest and the highest numbers stading for responses. Recoding the variables I want the livest number stand for most negative and the highest number stands for most positive meaning of a response.
- there is just one continuous variable that is agea, the rest of the variables are categorical

I  will display descriptive statistics by variable groups:
*General information on a participants*
```{r}
#load necessary libraries
library(ggplot2)
library(GGally)
library(haven)
library(tidyverse)
library(corrplot)
library(tidyverse)
library(dplyr)

#make one dataframe that contains all the cat variables. Dont do the same for age and gender since there is just one variable that is continuous or binary.
interested_var_cat <- select(data, "stflife", "polintr", "actrolga", "imwbcnt", "imbgeco", "imueclt", "imsmetn", "imdfetn", "impcntr")
# "gndr"
# agea

gather(interested_var_cat) %>% ggplot(aes(x = value)) + geom_bar(aes(y = value), col = "black", fill = "pink", alpha = 0.8, bins = 10)
```